---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# stressPLS

`stressPLS` is an R package scaffold for stress-testing formative
higher-order constructs in PLS-SEM. The package is being built around a clear
separation between model specification, perturbation generation, model
estimation, robustness scoring, and reporting.

The package does not fabricate PLS-SEM estimates. Instead, it creates validated
model and perturbation objects, runs user-supplied estimator backends, and
standardizes the resulting diagnostics. The examples below use a deterministic
toy backend that is only for demonstrating the API.

## Installation

```{r, eval = FALSE}
# install.packages("pak")
pak::pak("stressPLS")
```

## Minimal example

```{r}
library(stressPLS)

dat <- data.frame(
  x1 = rnorm(20),
  x2 = rnorm(20),
  x3 = rnorm(20),
  y = rnorm(20)
)

model <- list(
  indicators = c("x1", "x2", "x3"),
  paths = list(y = c("x1", "x2", "x3"))
)

grid <- stress_indicators(dat, model)
result <- stress_pls(dat, model, grid = grid)

summarise_stress(result)
rank_fragility(result)
```

## Model specification example

```{r}
image <- specify_construct(
  name = "Image",
  indicators = c("img1", "img2", "img3")
)

quality <- specify_construct(
  name = "Quality",
  indicators = c("qual1", "qual2", "qual3")
)

satisfaction <- specify_construct(
  name = "Satisfaction",
  indicators = c("sat1", "sat2", "sat3"),
  mode = "reflective"
)

brand_equity <- specify_hoc(
  name = "BrandEquity",
  dimensions = c("Image", "Quality")
)

paths <- specify_paths(
  from = "BrandEquity",
  to = "Satisfaction"
)

model_spec <- specify_model(
  constructs = list(image, quality, satisfaction),
  hocs = list(brand_equity),
  paths = paths
)

construct_names(model_spec)
required_indicators(model_spec)
```

## Perturbation example

```{r}
indicator_grid <- perturb_indicators(
  model = model_spec,
  method = "leave_one_out",
  constructs = c("Image", "Quality")
)

indicator_grid
```

## Custom backend example

Backends are regular R functions with the signature
`function(model, data, scenario = NULL, ...)`. They should return canonical
tables such as `paths`, `weights`, `vifs`, `r2`, and prediction metrics.

```{r}
demo_data <- data.frame(
  img1 = 1:12,
  img2 = 2:13,
  img3 = 3:14,
  qual1 = 4:15,
  qual2 = 5:16,
  qual3 = 6:17,
  sat1 = 7:18,
  sat2 = 8:19,
  sat3 = 9:20,
  group = rep(c("a", "b"), each = 6)
)

toy_backend <- function(model, data, scenario = NULL, ...) {
  brand_score <- rowMeans(data[c("img1", "img2", "img3", "qual1", "qual2", "qual3")])
  sat_score <- rowMeans(data[c("sat1", "sat2")])
  path_estimate <- unname(coef(lm(sat_score ~ brand_score))[2])

  list(
    paths = data.frame(
      from = "BrandEquity",
      to = "Satisfaction",
      estimate = path_estimate,
      significant = TRUE
    ),
    weights = data.frame(
      construct = "Image",
      indicator = c("img1", "img2", "img3"),
      estimate = colMeans(data[c("img1", "img2", "img3")]) / 100
    ),
    vifs = data.frame(
      construct = "Image",
      indicator = c("img1", "img2", "img3"),
      vif = c(1.2, 1.3, 1.4)
    ),
    prediction = data.frame(outcome = "sat1", metric = "rmse", value = 1)
  )
}

backend <- as_backend(toy_backend, name = "toy-example")
baseline <- fit_baseline_model(model_spec, demo_data, backend)
fit_grid <- fit_perturbation_grid(model_spec, demo_data, indicator_grid, backend)
```

## Stability metrics example

```{r}
calc_path_stability(fit_grid)
calc_indicator_fragility(fit_grid)
rank_fragility(fit_grid)
```

## Bootstrap example

```{r}
boot <- bootstrap_stability(model_spec, demo_data, backend, R = 5, seed = 1)
summarise_bootstrap_stability(boot)
```

## Collinearity stress example

```{r}
col_stress <- collinearity_stress_test(
  model_spec,
  demo_data,
  backend,
  indicators = c("img1", "img2", "img3"),
  levels = c(0, 0.3)
)
make_collinearity_table(col_stress)
```

## Prediction validation example

```{r}
cv_backend <- function(model, data, scenario = NULL, ...) {
  test <- scenario$test_data[[1]]
  list(predictions = data.frame(
    outcome = "sat1",
    observed = test$sat1,
    predicted = mean(data$sat1)
  ))
}

cv <- repeated_cv_predict(
  model_spec,
  demo_data,
  cv_backend,
  outcomes = "sat1",
  v = 3,
  repeats = 1,
  seed = 1
)
compare_prediction_metrics(cv)
```

## Heterogeneity example

```{r}
het <- subgroup_heterogeneity(
  model_spec,
  demo_data,
  backend,
  group = "group",
  min_n = 3
)
compare_subgroup_paths(het)
```

## Simulation example

```{r}
sim <- simulate_formative_hoc_data(n = 30, seed = 1)
simulation <- run_simulation_grid(data.frame(n = 30), backend, replications = 1, seed = 1)
summarise_simulation_results(simulation)
```

## Sensitivity report object

```{r}
report <- sensitivity_report(
  baseline = baseline,
  perturbations = fit_grid,
  bootstrap = boot,
  collinearity = col_stress,
  prediction = cv,
  heterogeneity = het,
  simulation = simulation,
  limitations = "Examples use a deterministic toy backend, not a PLS-SEM estimator."
)
report
```

## Concept

Stress testing asks whether substantive conclusions survive reasonable changes
to indicators, higher-order construct specifications, weights, bootstrap
settings, and related assumptions. `stressPLS` will expose those changes as
explicit scenario grids, run them through a modular backend, and report
stability and fragility without fabricating estimates when a backend is absent.
Future estimation backends will consume `stresspls_model_spec` objects and
`stresspls_perturbation_grid` objects.

## Current limitations

`stressPLS` is an experimental package. The safest supported route is a custom
backend supplied by the user. Adapter functions for `seminr`, `cSEM`, `plspm`,
and SmartPLS exports are documented skeletons unless users provide package-
specific estimator or parser functions. Simulation and stress diagnostics are
intended for robustness analysis, not as substitutes for a validated PLS-SEM
estimator.
