---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# stressPLS

`stressPLS` is an R package scaffold for stress-testing formative
higher-order constructs in PLS-SEM. The package is being built around a clear
separation between model specification, perturbation generation, model
estimation, robustness scoring, and reporting.

The initial package does not implement a PLS-SEM estimator. Instead, it creates
validated perturbation grids and result objects that can later be connected to
modular estimation backends.

## Installation

```{r, eval = FALSE}
# install.packages("pak")
pak::pak("stressPLS")
```

## Minimal example

```{r}
library(stressPLS)

dat <- data.frame(
  x1 = rnorm(20),
  x2 = rnorm(20),
  x3 = rnorm(20),
  y = rnorm(20)
)

model <- list(
  indicators = c("x1", "x2", "x3"),
  paths = list(y = c("x1", "x2", "x3"))
)

grid <- stress_indicators(dat, model)
result <- stress_pls(dat, model, grid = grid)

summarise_stress(result)
rank_fragility(result)
```

## Model specification example

```{r}
image <- specify_construct(
  name = "Image",
  indicators = c("img1", "img2", "img3")
)

quality <- specify_construct(
  name = "Quality",
  indicators = c("qual1", "qual2", "qual3")
)

satisfaction <- specify_construct(
  name = "Satisfaction",
  indicators = c("sat1", "sat2", "sat3"),
  mode = "reflective"
)

brand_equity <- specify_hoc(
  name = "BrandEquity",
  dimensions = c("Image", "Quality")
)

paths <- specify_paths(
  from = "BrandEquity",
  to = "Satisfaction"
)

model_spec <- specify_model(
  constructs = list(image, quality, satisfaction),
  hocs = list(brand_equity),
  paths = paths
)

construct_names(model_spec)
required_indicators(model_spec)
```

## Perturbation example

```{r}
indicator_grid <- perturb_indicators(
  model = model_spec,
  method = "leave_one_out",
  constructs = c("Image", "Quality")
)

indicator_grid
```

## Concept

Stress testing asks whether substantive conclusions survive reasonable changes
to indicators, higher-order construct specifications, weights, bootstrap
settings, and related assumptions. `stressPLS` will expose those changes as
explicit scenario grids, run them through a modular backend, and report
stability and fragility without fabricating estimates when a backend is absent.
Future estimation backends will consume `stresspls_model_spec` objects and
`stresspls_perturbation_grid` objects.
