---
title: "stressPLS methodology"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{stressPLS methodology}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(stressPLS)
```

## Overview

`stressPLS` is designed to evaluate whether formative higher-order construct
results in PLS-SEM remain stable under reasonable perturbations.

The framework separates five responsibilities:

1. Model specification.
2. Perturbation generation.
3. Model estimation.
4. Robustness scoring.
5. Reporting and visualization.

This initial scaffold implements the package structure, input validation,
perturbation grids, S3 result classes, summaries, plotting, and reporting
placeholders. It does not implement the full methodology or a PLS-SEM backend.

## Model specification objects

Model specifications are explicit declaration objects. Lower-order constructs
are created with `specify_construct()`, higher-order constructs are created
with `specify_hoc()`, structural paths are created with `specify_paths()`, and
the pieces are combined with `specify_model()`.

```{r}
image <- specify_construct("Image", c("img1", "img2", "img3"))
quality <- specify_construct("Quality", c("qual1", "qual2", "qual3"))
satisfaction <- specify_construct(
  "Satisfaction",
  c("sat1", "sat2", "sat3"),
  mode = "reflective"
)

brand_equity <- specify_hoc("BrandEquity", c("Image", "Quality"))
paths <- specify_paths(from = "BrandEquity", to = "Satisfaction")

model <- specify_model(
  constructs = list(image, quality, satisfaction),
  hocs = list(brand_equity),
  paths = paths
)

validate_model_spec(model)
construct_names(model)
required_indicators(model)
```

These objects do not estimate PLS-SEM models. They validate and preserve model
intent so later estimation backends can consume `stresspls_model_spec` objects
without needing to infer construct definitions from ad hoc lists.

## Perturbation families

The planned stress-test families include indicator deletion, indicator swapping,
construct reweighting, bootstrap sensitivity, and alternative higher-order
construct specifications. Current functions generate explicit scenario grids
that backends can consume later.
