---
title: "stressPLS methodology"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{stressPLS methodology}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(stressPLS)
```

## Overview

`stressPLS` is designed to evaluate whether formative higher-order construct
results in PLS-SEM remain stable under reasonable perturbations.

The resilience-analysis framing is deliberately conservative: the package asks
how much a conclusion changes when indicators, specifications, resampling
choices, collinearity pressure, prediction splits, or observed subgroups change.
It does not create estimates without an estimator backend.

The framework separates five responsibilities:

1. Model specification.
2. Perturbation generation.
3. Model estimation.
4. Robustness scoring.
5. Reporting and visualization.

This experimental release implements model declarations, perturbation grids,
backend standardization, diagnostic metrics, resampling and stress diagnostics,
summary tables, plotting, reports, and simulation utilities. It does not include
a validated built-in PLS-SEM estimator; real empirical use requires a
scientifically appropriate user-supplied backend or a future tested adapter.

The terminology is used consistently throughout the package:

- Robustness means conclusions remain substantively unchanged under plausible
  perturbations.
- Stability means weights, paths, signs, and predictive metrics remain
  consistent across resampling or perturbation.
- Resilience means the substantive interpretation withstands multiple
  stressors.
- Fragility means conclusions change materially under small or plausible
  changes.
- Predictive validity means satisfactory out-of-sample or cross-validation
  performance supplied or enabled by the backend.

## Model specification objects

Model specifications are explicit declaration objects. Lower-order constructs
are created with `specify_construct()`, higher-order constructs are created
with `specify_hoc()`, structural paths are created with `specify_paths()`, and
the pieces are combined with `specify_model()`.

```{r}
image <- specify_construct("Image", c("img1", "img2", "img3"))
quality <- specify_construct("Quality", c("qual1", "qual2", "qual3"))
satisfaction <- specify_construct(
  "Satisfaction",
  c("sat1", "sat2", "sat3"),
  mode = "reflective"
)

brand_equity <- specify_hoc("BrandEquity", c("Image", "Quality"))
paths <- specify_paths(from = "BrandEquity", to = "Satisfaction")

model <- specify_model(
  constructs = list(image, quality, satisfaction),
  hocs = list(brand_equity),
  paths = paths
)

validate_model_spec(model)
construct_names(model)
required_indicators(model)
```

These objects do not estimate PLS-SEM models. They validate and preserve model
intent so later estimation backends can consume `stresspls_model_spec` objects
without needing to infer construct definitions from ad hoc lists.

## Perturbation grid generation

Perturbation grids are explicit scenario tables generated from
`stresspls_model_spec` objects. The indicator perturbation layer can generate
leave-one-out deletion, random deletion, combinatorial deletion, replacement,
and noise-injection scenarios without estimating a model.

```{r}
indicator_grid <- perturb_indicators(
  model = model,
  method = "leave_one_out",
  constructs = c("Image", "Quality")
)

indicator_grid
```

Each scenario records the method, affected construct, removed indicators,
replacement metadata, noise recipes, validity flags, and a stable scenario ID.
When data are supplied to `noise_injection()`, transformed scenario data can be
materialized; otherwise the grid stores a reproducible noise recipe. These
perturbation grids are designed to be consumed by estimation backends, which run
each scenario and return standardized estimates for downstream robustness
metrics.

## Estimator backend abstraction

Backends are functions with signature `function(model, data, scenario = NULL,
...)`. A backend may wrap a PLS-SEM package, parse an export, or implement a
researcher-controlled estimator. `stressPLS` standardizes backend output into
tables for paths, weights, VIFs, R-squared values, prediction metrics,
diagnostics, and metadata.

The package includes optional adapter skeletons for common workflows, but custom
backends are currently the safest supported route. Skeleton adapters do not
imply tested support for those estimator packages until an estimator or parser
is supplied and validated by the user.

## Robustness metrics

The metric layer compares standardized baseline and scenario fits. It reports
sign consistency, direction flip rates, confidence interval width, scaled
interval width, stability indices, path stability, weight stability, and
indicator fragility. Metrics are computed from backend estimates only.

## Bootstrap stability

`bootstrap_stability()` resamples rows with replacement, fits each sample
through the backend, and summarizes sign consistency, interval width, stability
indices, convergence rates, and errors. Small bootstrap counts are useful for
examples; real studies should use larger `R` values.

## Collinearity stress

`inflate_collinearity()` approximates controlled collinearity pressure by
blending selected numeric indicators with a shared standardized component.
`collinearity_stress_test()` fits models across a grid of pressure levels and
tracks VIFs, path changes, weight changes, and available prediction metrics.
Exact target correlations are not guaranteed for arbitrary observed data.

## Predictive validation

`repeated_cv_predict()` creates repeated folds and fits through the backend.
If the backend returns predictions, `stressPLS` computes RMSE and MAE. If the
backend returns precomputed prediction metrics, those are collected directly.

## Heterogeneity diagnostics

Observed subgroup heterogeneity is handled with `subgroup_heterogeneity()`.
The function splits data by an observed grouping variable, fits each subgroup,
and compares path and weight estimates across groups. Latent class segmentation
is intentionally outside the current scope.

## Simulation design

`simulate_formative_hoc_data()` creates small formative higher-order composite
data with known parameters. `run_simulation_grid()` repeats simulations across
design rows and stores bias, RMSE, sign consistency, direction flip rates,
interval width, stability indices, and convergence/error rates when available.

## Interpretation guidance

Stress results should be interpreted as robustness evidence. Stable signs,
small perturbation changes, reasonable VIF behavior, and consistent prediction
metrics strengthen confidence in a conclusion. Large direction flips, strong
indicator fragility, subgroup instability, or prediction deterioration should
trigger model review rather than automatic rejection.

## Limitations and future work

The package is experimental. Its core tests use deterministic toy backends for
API verification, not PLS-SEM estimation. Real research use requires a backend
that returns scientifically meaningful estimates. Future work can add tested
adapters for specific PLS-SEM packages and richer report rendering.

## Perturbation families

The planned stress-test families include indicator deletion, indicator swapping,
construct reweighting, bootstrap sensitivity, and alternative higher-order
construct specifications. Current functions generate explicit scenario grids
that backends can consume later.
